---
title: "QuaranTeam2"
author: "Lukas Coffey, Drew Klaubert, David Parks, Arthur Presnetsov"
date: "20 April 2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(EpiModel)
```



Import the data.
```{r}
counties <- read.csv("EpiModel_Data.csv", stringsAsFactors = FALSE)
str(counties)
```



Create variables needed for the model.
```{r}
# the infection probability rate
inf_prob_rate <- 0.25

# the infected death rate
death_rate <- 0.0138

# the infected recovery rate
recovery_rate <- 1 / 24.7

# per county activity rate
# using the log base 10 of the per county population density as a proxy for 
# transmissive acts per person per time period
activity_rate <- log10(counties$pop_density)

# manually correct baltimore city's activity rate to be equal to the
# maximum of the other counties
activity_rate[3] <- sort(activity_rate, decreasing = TRUE)[2]
```



Compute the SIR model for each county WITHOUT intervention.
```{r}
# number of time steps
n_steps <- 90

# create an empty list to capture the models
sir_models_ni <- list()

# instantiate the models and store them in a list
index <- 1

for (pop in counties$county_population) {
  # assumes no births and no deaths from natural causes
  param <- param.dcm(inf.prob = inf_prob_rate,
                     act.rate = activity_rate[[index]],
                     rec.rate = recovery_rate,
                     a.rate = 0,
                     ds.rate = 0,
                     di.rate = death_rate,
                     dr.rate = 0)
  # assumes full population is susceptible, 1 infected person
  init <- init.dcm(s.num = pop - 1, i.num = 1, r.num = 0)
  control <- control.dcm(type = "SIR",
                         nsteps = n_steps)
  sir_model <- dcm(param, init, control)
  sir_models_ni[[index]] <- sir_model
  index <- index + 1
}
```



Compute the SIR model for each county WITH intervention starting on day 1, reducing the probability of transmission by 72%.
```{r}
effectiveness <- 0.72
start_day <- 1

# number of time steps
n_steps <- 90

# create an empty list to capture the models
sir_models_i <- list()

# instantiate the models and store them in a list
index <- 1

for (pop in counties$county_population) {
  # assumes no births and no deaths from natural causes
  param <- param.dcm(inf.prob = inf_prob_rate,
                     inter.eff = effectiveness,
                     inter.start = start_day,
                     act.rate = activity_rate[[index]],
                     rec.rate = recovery_rate,
                     a.rate = 0,
                     ds.rate = 0,
                     di.rate = death_rate,
                     dr.rate = 0)
  # assumes full population is susceptible, 1 infected person
  init <- init.dcm(s.num = pop - 1, i.num = 1, r.num = 0)
  control <- control.dcm(type = "SIR",
                         nsteps = n_steps)
  sir_model <- dcm(param, init, control)
  sir_models_i[[index]] <- sir_model
  index <- index + 1
}
```



Determine when the number of hospital beds would be fully occupied if 4% of those infected need hospitalization.
```{r}
# % of infected that need hospitalization
hospitalization_rate <- 0.04

# create empty arrays to capture the results
icu_no_i <- rep(NA, length(counties$name))
normal_no_i <- rep(NA, length(counties$name))
max_inf_no_i <- rep(NA, length(counties$name))
icu_i <- rep(NA, length(counties$name))
normal_i <- rep(NA, length(counties$name))
max_inf_i <- rep(NA, length(counties$name))


for (i in 1:length(counties$name)) {
  # NO INTERVENTION ***********************************************************
  # number of hospitalized patients per day
  infected <- as.data.frame(sir_models_ni[[i]])$i.num
  hospitalized <- floor(infected * hospitalization_rate)
  total_beds <- counties$icu_bed_count[i] + counties$inpatient_bed_count[i]
  
  # get the day when the number of infected needing hospitalization
  # exceeds icu capacity
  icu_no_i[i] <- which.max(hospitalized > counties$icu_bed_count[i])
  
  # get the day when the number of infected needing hospitalization
  # exceeds total bed capacity
  normal_no_i[i] <- which.max(hospitalized > total_beds)
  
  # get the maximum number of infected
  max_inf_no_i[i] <- round(max(infected))
  # NO INTERVENTION ***********************************************************
  # ***************************************************************************
  # INTERVENTION **************************************************************
  # number of hospitalized patients per day
  infected <- as.data.frame(sir_models_i[[i]])$i.num
  hospitalized <- floor(infected * hospitalization_rate)
  
  # get the day when the number of infected needing hospitalization
  # exceeds icu capacity
  icu_i[i] <- which.max(hospitalized > counties$icu_bed_count[i])
  
  # get the day when the number of infected needing hospitalization
  # exceeds total bed capacity
  normal_i[i] <- which.max(hospitalized > total_beds)
  
  # get the maximum number of infected
  max_inf_i[i] <- round(max(infected))
  # INTERVENTION **************************************************************
}

# manually correct for counties without hospital beds
icu_no_i[counties$icu_bed_count == 0] <- 0
normal_no_i[counties$inpatient_bed_count == 0] <- 0
icu_i[counties$icu_bed_count == 0] <- 0
normal_i[counties$inpatient_bed_count == 0] <- 0

# create a table of the results
results <- cbind(counties$name,
                 icu_no_i,
                 normal_no_i,
                 max_inf_no_i,
                 icu_i,
                 normal_i,
                 max_inf_i)
colnames(results) <- c("County",
                       "ICU-No I",
                       "IPB-No I",
                       "MaxInf-No I",
                       "ICU-I",
                       "IPB-I",
                       "Max Inf-I")
print(as.table(results))
```



Compute the Root Sum Squared Error for the intervention model on day 45.
```{r}
# create an empty array to capture the results
estimated_cases <- rep(NA, length(counties$name))

# get the number of infected for each county on day 45
for (i in 1:length(counties$name)) {
  estimated_cases[[i]] <- round(as.data.frame(sir_models_i[[i]])$i.num[45])
}

# actual cases in MD on day 45
actual_cases <- c(33,1005,1378,1664,109,33,308,131,347,20,557,4,195,508,16,
                  2404,3160,24,101,10,16,116,138,31)

# create a comparison table
results <- cbind(counties$name,
                 actual_cases,
                 estimated_cases,
                 actual_cases - estimated_cases)
colnames(results) <- c("County",
                       "Actual Cases",
                       "Model Cases",
                       "Difference")
print(as.table(results))

# RSSE
sqrt(sum((actual_cases - estimated_cases)^2))
```



Plot the results
```{r}
for (i in 1:length(counties$name)) {
  # NO INTERVENTION *********************************************************
  # plot the results of the SIR model
  plot(sir_models_ni[[i]],
       main = paste(counties$name[i], "w\\o Intervention"),
       xlab = "Days", ylab = "Population",
       xlim = c(1, n_steps),
       ylim = c(0, counties$county_population[i] * 1.1))
  abline(v=45)
  
  
  # plot the number of daily deaths
  plot(sir_models_ni[[i]], y = "di.flow", lwd = 4, col = "firebrick",
       main = paste(counties$name[i], "Daily Deaths w\\o Intervention"))
  abline(v=45)
  # NO INTERVENTION *********************************************************
  # *************************************************************************
  # INTERVENTION ************************************************************
  # plot the results of the SIR model
  plot(sir_models_i[[i]],
       main = paste(counties$name[i], "with Intervention"),
       xlab = "Days", ylab = "Population",
       xlim = c(1, n_steps),
       ylim = c(0, counties$county_population[i] * 1.1))
  abline(v=45)
  
  # plot the number of daily deaths
  plot(sir_models_i[[i]], y = "di.flow", lwd = 4, col = "firebrick",
       main = paste(counties$name[i], "Daily Deaths with Intervention"))
  abline(v=45)
  # INTERVENTION ************************************************************
}
```